---
title: "My first markdown file"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
library(dplyr)
library(readr)
library(ggplot2)
```
## Reading the Data
## Plotting the Data

This data displays the favorite restaurants of the class. It includes the student's name, restaurant, location, cuisine, dish, cost per person, times visited per year, and last visit.
```{r}
mine <- read_csv("https://raw.githubusercontent.com/URstats/stat209-s18-laurenblose/master/fav_restaurants.csv")
```
This graph indicates the popularity of cuisines among the class data for favorite restaurants. 
```{r}
ggplot(mine, aes(cuisine)) +
  geom_bar()
```

This graph shows the cost per person at the favorite restaurants of the class in the data.
```{r}
ggplot(mine, aes(cost_per_person)) +
  geom_bar()
```

This graph shows the yearly visits to favorite restaurants compared to the cost per person.
```{r}
ggplot(mine, aes(yearly_visits, cost_per_person)) +
  geom_point()
```







