---
title: "2017 SEC vs PAC 12 Softball"
author: "Lauren Blose"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(out.width = "100%")
knitr::opts_chunk$set(dpi = 300)

library(readr)
library(ggplot2)
library(dplyr)
library(ggmap)
library(viridis)
library(forcats)
library(smodels)
```

```{r}
softball <- read_csv("https://raw.githubusercontent.com/URstats/stat209-s18-laurenblose/master/project_c.csv")
```

## Introduction

For my project, I chose to look deeper into some softball statistics from the 2017 season. I chose players from 4 teams from the the Southeastern Conference (SEC): LSU, Auburn, Arkansas, Florida, Alabama, and Kentucky. I also chose players from 4 teams from the Pacific-12 Conference (PAC-12): Oregon, Arizona State, Oregon State, Stanford, UCLA and Cal. I am interested in looking at the batting averages and slugging percentages of the players, and overall as the SEC and PAC-12. My hypotheses are as follows:

1. I hypothesize that the SEC will have a higher overall batting average than the PAC-12.
2. I hypothesize that the SEC will have a higher overall slugging percentage than the PAC-12.
3. I hypothesize that there will be a significant and positive relationship between batting averages and slugging percentages.

## Methods

In order to collect my data, I went to the team website for each of the teams that I decided to use in my project. I recorded the batting average and slugging percentage for 8 players of each of the teams that listed earlier. The batting average of each player is a ratio of a batter's safe hits per official at bats, and slugging percentage is the total number of bases covered by a player divided by the number of at bats. These were calculated by the teams and were published online. For my first 2 tests to explore my first 2 hypotheses, I used boxplots to compare batting averages and slugging percentages to the conferences. For my third test, I used a scatterplot with a linear regression for batting average vs slugging percentage. The baseline group in the first 2 regression tables is the PAC-12 and that of the third test is slugging percentage.

## Results

Show the output of the inference and regression analysis. Along with each model, include a graphic that describes the patterns shown in the model. Make sure you use full sentences to describe each output, but there is no need to write too much in this section.

```{r}
ggplot(softball, aes(conference, batting_avg)) +
  geom_boxplot() +
  xlab("Conference") +
  ylab("Batting Average") +
  ggtitle("Conference and Batting Average")
```

```{r}
model <- lm_basic(batting_avg ~ 1 + conference, data = softball)
reg_table(model, level = 0.95)
```

The above plot and regression table indicate that batting average does not differ significantly between the SEC and PAC-12.

```{r}
ggplot(softball, aes(conference, slugging_perc)) +
  geom_boxplot() +
    xlab("Conference") +
  ylab("Slugging Percentage") +
  ggtitle("Conference and Slugging Percentage")
```

```{r}
model <- lm_basic(slugging_perc ~ 1 + conference, data = softball)
reg_table(model, level = 0.95)
```

The above plot and regression table indicate that slugging percentage does not differ significantly between the SEC and PAC-12.

```{r}
qplot(batting_avg, slugging_perc, data = softball) + geom_smooth(method = "lm") +
    ylab("Slugging Percentage") +
  xlab("Batting Average") +
  ggtitle("Slugging Percentage and Batting Average")
```

```{r}
model <- lm_basic(slugging_perc ~ 1 + batting_avg, data = softball)
reg_table(model, level = 0.95)
```

This test indicates a significant and positive relationship between batting average and slugging percentage in that via the formula: (batting average x 1.000917) + .157689. Because the slope is so close to one, it indicates that batting average and slugging percentage increase together. 

## Conclusions

Test 1, which compared the batting averages of the SEC and PAC-12 to one another, resulted in a slightly negative slope. However, because this value was -0.009042 and the other values in the regression table were negative and positive, the difference between the batting averages of the SEC and PAC-12 are insignificant. Test 2 compared the slugging percentages of the SEC and PAC-12 and resulted in a positive slope of 0.008417. Similarly to Test 1, the other values in the regression table were negative and positive, making the difference between slugging percentages of the SEC and PAC-12 insignificantly different. Test 3, as I predicted, rendered a significant and positive relationship between batting average and slugging percentage in that slugging percentage tends to generally increase with batting average. 

## References

http://www.lsusports.net/src/data/lsu/assets/docs/sb/17stats/teamcume.htm?DB_OEM_ID=5200

http://www.auburntigers.com/sports/w-softbl/stats/2016-2017/teamcume.html

http://floridagators.com/documents/2017/6/20//2017_SB_Overall_Final.pdf?id=18080

http://www.arkansasrazorbacks.com/wp-content/uploads/2017/05/2017-ARKSB-Season-Statistics_0520.pdf

https://thesundevils.com/documents/2017/3/28/Overall_Season.pdf?path=softball

http://goducks.com/cumestats.aspx?path=softball&year=2017

http://gostanford.com/cumestats.aspx?path=softball&year=2017

https://static.osubeavers.com/custompages/SOFTBALL/SB2017/HTML/teamcume.htm

https://s3.amazonaws.com/rolltide.com/documents/2017/2/10/season_stats.pdf

http://ukathletics.com/sports/2017/11/9/2017-softball-stats.aspx

http://static.pac-12.com/sports/softball/stats/2016-17/HTML/ucla.htm#team.mlb

http://static.pac-12.com/sports/softball/stats/2016-17/HTML/cal.htm#team.mlb

